---
title: "Activity Four - Lab1"
author: "Taehoon Ha"
date: "10/04/2018"
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Write a function that normalizes a vector (with two parameters):  
  * vector to be normalized.
  * parameter that allows normalize the vector two ways: only centered with zero mean and also standardize with mean zero and standard deviation 1.
  
Example:    
  * input(1, 2, 3, 4)
  * centered vector is
```{r}
standardize <- function (vec, centered_only = F) {
  
  # create an empty vector in order to save the output result
  result <- c()
  
  # centered_only = F
  if (!centered_only) {
    for (i in 1:length(vec)) {
      result[i] <- (vec[i] - mean(vec)) / sd(vec) 
    }
  
  # centered_only = T
  } else {
    for (i in 1:length(vec)) {
      result[i] <- vec[i] - mean(vec) 
    }
  }
  
  # return the result vector
  return(result)  
}

a <- c(1, 2, 3, 4)
standardize(a, centered_only = T)
standardize(a, centered_only = F)
```

### 2. Modify previous function in order to handle missing values in a vector.
```{r}
standardize <- function (vec, centered_only=F, na.rm=F) {
  
  # create an empty vector in order to save the output result
  result <- c()
  
  # 1. centered_only = F
  if(!centered_only) {
    
    # 1.a centered_only = F, na.rm = F
    if(!na.rm) {
      for (i in 1:length(vec)) {
        result[i] <- (vec[i] - mean(vec)) / sd(vec)}}
    
    # 1.b centered_only = F, na.rm = T
    else {
      for (i in 1:length(vec)) {
        result[i] <- (vec[i] - mean(vec, na.rm = T)) / sd(vec, na.rm = T)
      }
    }
  }
  
  # 2. centered_only = F
  else {
    
    # 2.a centered_only = T, na.rm = F
    if(!na.rm) {
      for (i in 1:length(vec)) {
        result[i] <- vec[i] - mean(vec)}}
    
    # 2.b centered_only = T, na.rm = T
    else {
      for (i in 1:length(vec)) {
        result[i] <- vec[i] - mean(vec, na.rm = T)
      }
    }
  }
  
  # return the result
  return(result)
}

standardize(c(1,2,3,4,NA))
standardize(c(1,2,3,4,NA), na.rm = T)
```

### 3. Create a 4 by 4 matrix
Create a function to calculate sum of matrix elements without built-in function sum()
```{r}

# The function below automatically generates 4 by 4 table filled with 1:16 vector and calculated the sum of matrix.

matrix_sum <- function() {
  
  # generating a 4 by 4 table filled with the given vector, 1:16
  mat <- matrix(1:16, 4, 4)
  
  # the vector for saving the sum
  mat.sum <- 0
  
  # row-indexing from 1 to 4
  for (i in 1:nrow(mat)) {
    
    # col-indexing begin from 1 to 4
    for (j in 1:ncol(mat)) {
      
      # saving the sum result in the vector
      mat.sum <- mat.sum + mat[i, j]
    }
  }
  
  # show the matrix used and return the sum of the matrix
  print(mat)
  return(mat.sum)
}

matrix_sum()
```


```{r}

# Advanced

matrix_sum <- function (vec, nrow, ncol) {
  mat <- matrix(vec, nrow = nrow, ncol = ncol)
  mat.sum <- 0
  
  for (i in 1:nrow(mat)) {
    for (j in 1:ncol(mat)) {
      mat.sum <- mat.sum + mat[i, j]
    }
  }
  print(mat)
  cat(paste0("\nThis is ", nrow, " by ", ncol, " matrix.\n", "The sum of matrix is: \n"))
  return(mat.sum)
}

matrix_sum(vec = 1:16, nrow = 4, ncol = 4)

# Other Example
matrix_sum(vec = 1:100, nrow = 20, ncol = 5)
```
